---
title: "Coursera Assignment"
author: "Sylva"
date: "2024-05-20"
output:
  html_document: default
  pdf_document: default
---
# Preparations 
```{r}
library(ggplot2)
library(dplyr)
library(lattice)
```

```{r setup, include=TRUE}
knitr::opts_chunk$set(# tidy.opts=list(width.cutoff=80),tidy=TRUE,
        dev = c("png"), 
        dpi = 300, 
        fig.show = "hold",
        fig.align = "center")
```

# Reading in the activity data
```{r}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```
# What is the mean number of steps taken each day?
## Histogram of steps taken per day 
```{r histogram of steps per day}
daily_steps <- aggregate(steps ~ date, data = activity, sum)
ggplot(daily_steps, aes(x = steps)) +
  geom_histogram(binwidth = 50, fill = 'blue', color = 'black', alpha = 0.7) +
  labs(title = 'Histogram of Steps Taken Each Day',
       x = 'Number of Steps',
       y = 'Frequency') +
  theme_light() +
  scale_x_continuous(breaks = seq(0, max(daily_steps$steps) + 50, by = 2000))
```

## Calculate the mean and median number of steps taken per day
```{r}
mean(daily_steps$steps)
median(daily_steps$steps)
```

# What is the average daily activity pattern?
## Time series plot of 5 minute intervals 

```{r time series plot of 5 minute intervals}
# Remove rows with NA in 'steps'
activity_no_na <- activity[!is.na(activity$steps), ]

# Group by interval and calculate the average steps
interval_avg <- aggregate(steps ~ interval, data = activity_no_na, FUN = mean)

plot(interval_avg$interval, interval_avg$steps, type = "l", col = "blue", pch = 16, 
     xlab = "Interval (minutes)", ylab = "Average Number of Steps", 
     main = "Average Number of Steps per 5-Minute Interval")
grid()
```
# Sum of NA in total 
```{r}
sum_na <- sum(is.na(activity$steps))
```

# Strategy for imputation
```{r}
# Calculation the mean for each 5-minute interval
interval_means <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)

# Create a named vector of interval means
interval_mean_vector <- setNames(interval_means$steps, interval_means$interval)

# Function to impute missing values with the mean for the respective interval
impute_steps <- function(steps, interval) {
  if (is.na(steps)) {
    return(interval_mean_vector[as.character(interval)])
  } else {
    return(steps)
  }
}

activity_imp <- activity
# Apply the imputation function to a new dataframe
activity_imp$steps <- mapply(impute_steps, activity_imp$steps, activity_imp$interval)

# Verify the imputation
head(activity_imp, 20)

# Histogram of total steps taken per day

daily_steps <- aggregate(steps ~ date, data = activity_imp, sum)
ggplot(daily_steps, aes(x = steps)) +
  geom_histogram(binwidth = 50, fill = 'blue', color = 'black', alpha = 0.7) +
  labs(title = 'Histogram of Steps Taken Each Day',
       x = 'Number of Steps',
       y = 'Frequency') +
  theme_light() +
  scale_x_continuous(breaks = seq(0, max(daily_steps$steps) + 50, by = 2000))

mean(daily_steps$steps)
median(daily_steps$steps)
```

# Calculation weekdays or weekends
```{r}
# Create the dayofweek variable
activity_imp$dayofweek <- ifelse(weekdays(activity_imp$date) %in% c("Samstag", "Sonntag"), "Weekend", "Weekday")

# Convert dayofweek to a factor
activity_imp$dayofweek <- factor(activity_imp$dayofweek, levels = c("Weekday", "Weekend"))
```

# Make a time series plot for weekdays and weekends
```{r time series of weekdays and weekends}
# Calculate the average steps for each interval and day type
interval_avg <- aggregate(steps ~ interval + dayofweek, data = activity_imp, FUN = mean)

# Create the panel plot using lattice
xyplot(steps ~ interval | dayofweek, data = interval_avg, type = "l", layout = c(1, 2),
       xlab = "Interval (minutes)", ylab = "Average Number of Steps",
       main = "Average Number of Steps per 5-Minute Interval by Day Type",
       col.line = "blue", lwd = 1)
```

